---
title: "Write a API client with apihelpers"
author: "Mahmoud Shaaban"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Write a API client with apihelpers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Overview

Provide functions for constructing a URL, sending requests and capturing API 
responses. The query parameters are checked and formatted according to a 
`query_map` which describes the type of requests and queries allowed by the 
API.

## Installation

Install `apihelpers` from GitHub using `devtools`

```r
devtools::install_github('abifromr/apihelpers')
```

## What `apihelpers` does?

This function should be able to do three things:
  1. Take the query paramters as an input from the user
  2. Send an API request
  3. Return the API response in an approptiate format

`apihelpers` provide functions to help with these three steps; `make_query`,
`send_request` and `format_content`.

Because Each API is different, we use a `query_map` which describe and document
the query paramters specific to the interface. Included in the package two 
datasets called `query_map_string` and `query_map_bioware` which describe the 
STRING (v11) and BIOWARE APIs.

## Example from `stringapi`

```{r}
# loading required libraries
library(apihelpers)
library(httr)
```

```{r}
# show query_map for string api
query_map_string
```

```{r}
# write a wrapper function
get_string_ids <- function(identifiers = NULL, species = NULL) {
  # construct query
  param <- list(identifiers = identifiers,
                species = species)
  
  query <- make_query(request = 'get_string_ids',
                      parameters = param,
                      query_map = query_map_string)
  
  # make url
  url <- modify_url('',
                    scheme = 'https',
                    hostname = 'string-db.org',
                    path = 'api/tsv/get_string_ids',
                    query = query)

  # get response
  resp <- send_request(url)

  # format contents
  res <- format_content(resp)

  # return results
  return(res)
}
```

```{r}
# make the call to api
res <- get_string_ids(identifiers = c('p53', 'dcdk2'),
                      species = 9606)
```

```{r}
# show results of the api call
head(res)
```

